---
title: "OxyProxy: Vectorized Inputs Workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{OxyProxy: Vectorized Inputs Workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE
)
```

## Overview

This vignette demonstrates how to run oxyproxy using **vector inputs** to explore multiple scenarios at once.

Common use cases:
- Sensitivity to air temperature and humidity
- Differences in surface water δ18O across sites
- Comparing diets or species masses across populations

## Load the package

```{r}
library(oxyproxy)
```

## Step 1 — Provide vector environmental inputs

`environment_function()` accepts vectors and returns a **full factorial** design.

```{r}
env_grid <- environment_function(
  air_temperature = c(10, 20),
  relative_humidity = c(0.50, 0.75),
  d18O_surface_water = c(-2, -5)
)

env_grid
nrow(env_grid)
```

### What to expect

The number of rows equals:

`length(air_temperature) * length(relative_humidity) * length(d18O_surface_water)`

In this example: `2 * 2 * 2 = 8` rows.

## Step 2 — Hold species and food constant, vary environment via the wrapper

This is a typical sensitivity analysis: one species + one diet, many environments, run in a single
call to `oxy_proxy_function()`.

```{r}
oxyproxy_many <- oxy_proxy_function(
  model_bodymass = 70,
  model_WaterEconomyIndex = 0.40,
  model_digestibility_of_food = 0.8,
  model_Carbohydrate_Content = 0.5,
  model_Protein_Content = 0.2,
  model_Fat_Content = 0.3,
  model_Free_Water_Content_Food = 0.7,
  model_air_temperature = c(10, 20),
  model_Relative_Humidity = c(0.50, 0.75),
  model_d18O_surfacewater = c(-2, -5),
  changeConstant = FALSE,
  SweatingSpecies = FALSE,
  PlotRange = FALSE   # set TRUE if you want a plot for visualization
)

oxyproxy_many
```

## Step 3 — Explicit scenario design across species, diet, environment

If you want to vary *everything* (e.g., body mass AND temperature AND diet), build a scenario table
and run row-wise using the individual pipeline functions.

```{r}
scenarios <- expand.grid(
  body_mass = c(20, 70),
  air_temperature = c(10, 20),
  relative_humidity = c(0.50, 0.75),
  d18O_surface_water = c(-2, -5),
  stringsAsFactors = FALSE
)

head(scenarios)
nrow(scenarios)
```

### Execute scenarios

```{r}
run_scenario <- function(row) {
  sp <- species_function(body_mass = row$body_mass, water_economy_index = 0.1)

  fd <- food_function(
    digestibility_of_food = 0.1,
    Carbohydrate_Content = 0.1,
    Protein_Content = 0.2,
    Fat_Content = 0.7,
    Free_Water_Content_Food = 0.1,
    changeConstant = FALSE
  )

  env <- environment_function(
    air_temperature = row$air_temperature,
    relative_humidity = row$relative_humidity,
    d18O_surface_water = row$d18O_surface_water
  )

  inputs_out <- input_function(species = sp, food = fd, environment = env)
  outputs_out <- outputs_function(inputs = inputs_out, sweating_species = FALSE)
  d18_obw_function(outputs = outputs_out)
}

outs <- lapply(seq_len(nrow(scenarios)), function(i) run_scenario(scenarios[i, , drop = FALSE]))
length(outs)
```

## Summarize results into a data frame

```{r}
extract_one <- function(out) {
  data.frame(d18Obw = out$d18Obw, stringsAsFactors = FALSE)
}

results_df <- do.call(rbind, lapply(outs, extract_one))
results_df <- cbind(scenarios, results_df)

head(results_df)
```

## Session info

```{r}
sessionInfo()
```
