---
title: "OxyProxy: Single-Value Workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{OxyProxy: Single-Value Workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE
)
```

## Overview

This vignette demonstrates a **single-run** workflow for predicting oxygen isotope ratios (δ18O) using oxyproxy, both step by step through the individual pipeline functions and via the `oxy_proxy_function()` wrapper.

## Load the package

```{r}
library(oxyproxy)

# If you're developing locally and not installed:
# devtools::load_all()
```

## Step 1 — Species physiology (single value)

`species_function()` expects `body_mass` and `water_economy_index` (plus optional `changeConstant`).

```{r}
species_out <- species_function(
  body_mass = 70,             # kg
  water_economy_index = 0.40, # typical range ~0.05-0.6
  changeConstant = FALSE
)

str(species_out)
```

## Step 2 — Diet composition (single value)

`food_function()` expects:
`digestibility_of_food`, `Carbohydrate_Content`, `Protein_Content`, `Fat_Content`, and `Free_Water_Content_Food`.

```{r}
food_out <- food_function(
  digestibility_of_food = 0.8,
  Carbohydrate_Content = 0.5,
  Protein_Content      = 0.2,
  Fat_Content          = 0.3,   # NOTE: must sum with carb+protein to 1.0
  Free_Water_Content_Food = 0.7,
  changeConstant = FALSE
)

str(food_out)
```

## Step 3 — Environment (single value)

`environment_function()` expects `air_temperature`, `relative_humidity`, and `d18O_surface_water`.

```{r}
env_out <- environment_function(
  air_temperature = 20,       # Celsius
  relative_humidity = 0.60,   # proportion in (0, 1]
  d18O_surface_water = -5     # per mil (‰ VSMOW)
)

env_out
```

## Step 4 — Inputs -> Outputs -> d18O body water

### 4A) Oxygen inputs

`input_function()` combines species, food, and environment into oxygen input variables:

```{r}
inputs_out <- input_function(
  species = species_out,
  food = food_out,
  environment = env_out
)

str(inputs_out)
```

### 4B) Oxygen outputs

`outputs_function()` computes the oxygen output variables:

```{r}
outputs_out <- outputs_function(
  inputs = inputs_out,
  sweating_species = FALSE
)

str(outputs_out)
```

### 4C) Final isotope estimates

`d18_obw_function()` computes the final body water, phosphate, and carbonate isotope estimates:

```{r}
d18o_out <- d18_obw_function(outputs = outputs_out)

head(d18o_out[, c("d18Obw", "d18Ophos", "d18Ocarb")])
```

## Alternative — Call the wrapper `oxy_proxy_function()`

`oxy_proxy_function()` runs the full pipeline in one call. It does **not** accept `species=`, `food=`, `environment=`;
instead it accepts a set of `model_*` scalar inputs:

```{r}
oxyproxy_out <- oxy_proxy_function(
  model_bodymass = 70,
  model_WaterEconomyIndex = 0.40,
  model_Digestibility_of_food = 0.8,
  model_Carbohydrate_Content = 0.5,
  model_Protein_Content = 0.2,
  model_Fat_Content = 0.3,
  model_Free_Water_Content_Food = 0.7,
  model_air_temperature = 20,
  model_Relative_Humidity = 0.60,
  model_d18O_surfacewater = -5,
  changeConstant = FALSE,
  sweating_species = FALSE,
  PlotRange = FALSE   # avoid plots in vignette by default
)

head(oxyproxy_out[, c("d18Obw", "d18Ophos", "d18Ocarb")])
```

## Inspecting results

```{r}
names(d18o_out)
str(d18o_out, max.level = 1)
```

## Session info

```{r}
sessionInfo()
```
